graphsim: An R package for simulating gene expression data from graph structures of biological pathways

### 2020-07-09 paper/paper.Rmd
copyediting paper https://github.com/openjournals/joss-reviews/issues/2161

### Summary
Transcriptomic analysis is used to capture the molecular state of a cell
or sample in many biological and medical applications. In addition to 
identifying alterations in activity at the level of individual genes, 
understanding changes in the gene networks that regulate fundamental
biological mechanisms is also an important objective of molecular 
analysis. As a result, databases that describe biological pathways 
are increasingly uesad to assist with the interpretation of results
from large-scale genomics studies. Incorporating information from 
biological pathways and gene regulatory networks into a genomic data
analysis is a popular strategy, and there are many methods that provide
this functionality for gene expression data. When developing or comparing
such methods, it is important to gain an accurate assessment of their 
performance. Simulation-based validation studies are frequently used
for this.
This necessitates the use of simulated data that correctly accounts for
pathway relationships and correlations. Here we present a versatile
statistical framework to simulate correlated gene expression data from
biological pathways, by sampling from a multivariate normal distribution
derived from a graph structure. This procedure has been released as the
\texttt{graphsim} R package on CRAN and GitHub (\url{https://github.com/TomKellyGenetics/graphsim})
 and is compatible with any graph structure that can be described using
the `igraph` package. This package allows the simulation of biological
pathways from a graph structure based on a statistical model of gene expression.

>>>>

### Summary
Transcriptomic analysis is used to capture the molecular state of a cell
or sample in many biological and medical applications. In addition to 
identifying alterations in activity at the level of individual genes, 
understanding changes in the gene networks that regulate fundamental
biological mechanisms is also an important objective of molecular 
analysis. As a result, databases that describe biological pathways 
are increasingly used to assist with the interpretation of results
from large-scale genomics studies. Incorporating information from 
biological pathways and gene regulatory networks into a genomic data
analysis is a popular strategy, and there are many methods that provide
this functionality for gene expression data. When developing or comparing
such methods, it is important to gain an accurate assessment of their 
performance. Simulation-based validation studies are frequently used
for this.
This necessitates the use of simulated data that correctly accounts for
pathway relationships and correlations. Here we present a versatile
statistical framework to simulate correlated gene expression data from
biological pathways, by sampling from a multivariate normal distribution
derived from a graph structure. This procedure has been released as the
\texttt{graphsim} R package on CRAN and GitHub (\url{https://github.com/TomKellyGenetics/graphsim})
 and is compatible with any graph structure that can be described using
the `igraph` package. This package allows the simulation of biological
pathways from a graph structure based on a statistical model of gene expression.
